---
title: "Analysis of Drone Data of Small Elephant Impact Sites"
author: "Sunniva McKeever, Isabella Metz and Maximilian Merzdorf"
date: "Internship Kruger WS23/24"
output: pdf_document
---


## Data Import

Load the required libraries.

```{r libraries, warning=FALSE, message=FALSE}
library(terra)
library(ggplot2)
library(lidR)
library(mapview)
library(sf)
library(knitr)
```

Set working directory and hyperparameters and import drone data.

```{r data import, warning=FALSE, message=FALSE}
setwd("C://Users/avinn/Documents/Master/Semester3/ElephantTransects/")

## (only once!!!) create empty data frame
params_df <- data.frame(matrix(ncol = 7, nrow = 0))
colnames(params_df) <- c("name", 
                         "numtrees", 
                         "treedens", 
                         "treeheight_min", 
                         "treeheight_max", 
                         "treeheight_mean", 
                         "canopyarea"
                         )

params_df_indices <- data.frame(matrix(ncol = 4, nrow = 0))
colnames(params_df_indices) <- c("ndvi_mean", 
                                 "evi_mean",
                                 "gci_mean",
                                 "lai_mean")


# hyperparameters
mw_size <- 15
crs_epsg <- "epsg:32736"
area <- 0.11 # area of each EIA in km2
EIA_name <- "EIA2 C3"

# load drone data
# EIA2C3
DSM <- terra::rast("./ElephantTransectSites/Pix4d/20230810_EIA2_C3/20230810_EIA2C3_dsm.tif")
DTM <- terra::rast("./ElephantTransectSites/Pix4d/20230810_EIA2_C3/20230810_EIA2C3_dtm.tif")
Ortho <- terra::rast(c(
  "./ElephantTransectSites/Pix4d/20230810_EIA2_C3/20230810_EIA2C3_transparent_mosaic_group1.tif",
  "./ElephantTransectSites/Pix4d/20230810_EIA2_C3/20230810_EIA2C3_transparent_mosaic_green.tif",
  "./ElephantTransectSites/Pix4d/20230810_EIA2_C3/20230810_EIA2C3_transparent_mosaic_red.tif",
  "./ElephantTransectSites/Pix4d/20230810_EIA2_C3/20230810_EIA2C3_transparent_mosaic_red edge.tif",
  "./ElephantTransectSites/Pix4d/20230810_EIA2_C3/20230810_EIA2C3_transparent_mosaic_nir.tif"),
  lyrs = c(1,2,3,5,7,9,11)
)

# rename bands
names(Ortho) <- c("red", "green", "blue", "MSgreen", "MSred", "MSrededge", "MSnir")

# load other data
aoi <- st_read("./data/other/polygons.shp")
```


## Data Preprocessing

The data needs to be reprojected (as lidR package requires a projection in meters) and cropped to the extent of the EIA. Then, an absolute tree height raster is calculated by subtracting the DTM from the DSM.

```{r data preprocessing,  warning = FALSE, message=FALSE}
## DATA PREPROCESSING
# reproject
# lidR package requires projection in m
DSM <- terra::project(DSM, crs_epsg)
DTM <- terra::project(DTM, crs_epsg)
aoi <- sf::st_transform(aoi, crs = crs_epsg)

# crop
DSM <- terra::mask(DSM, aoi[aoi$FieldID == EIA_name,])
DTM <- terra::mask(DTM, aoi[aoi$FieldID == EIA_name,])
Ortho <- terra::mask(Ortho, aoi[aoi$FieldID == EIA_name,])

# calculate Canopy Height Model (CHM) from DSM and DTM
DSM <- resample(DSM, DTM)
CHM <- DSM - DTM
CHM <- aggregate(CHM, 10) # lower resolution to limit computational time

# plot
par(mfrow = c(1,3))
plot(DSM, main = "DSM")
plot(DTM, main = "DTM")
plot(CHM, main = "CHM")
```

## Analysis
Basic indices are calculated to compare the structure of different EIAs among each other. Using the lidR package, individual trees are then detected and segmented. Tree tops can be detected by applying a Local Maximum Filter (LMF) on the loaded data set.For a given point, the algorithm analyzes neighborhood points, checking if the processed point is the highest. The size of the moving window determines the size of the analysed neighborhood.

```{r analysis,  warning=FALSE, include=TRUE, message=FALSE}
# calculate indices
ndvi <- (Ortho$MSnir - Ortho$red) / (Ortho$MSnir + Ortho$red)
evi <- 2.5 * ((Ortho$MSnir - Ortho$red) / (Ortho$MSnir + 6 * Ortho$red - 7.5 * Ortho$blue + 1))
gci <- (Ortho$MSnir / Ortho$MSgreen) - 1
lai <- 3.618 * evi - 0.118

# locate tree tops
# a tree in savannah is everything > 1.5m
ttops <- locate_trees(CHM, lmf(ws = mw_size, hmin = 1.5)) 

# segment trees
algo <- lidR::dalponte2016(CHM, ttops)
crowns <- algo()

# calculate parameters
numtrees <- round(nrow(ttops), digits = 2)
treedens <- round(numtrees/area, digits = 2) # number of trees per km2
treeheight_min <- round(min(ttops$Z), digits = 2)
treeheight_max <- round(max(ttops$Z), digits = 2)
treeheight_mean <- round(mean(ttops$Z), digits = 2)
canopyarea <- terra::expanse(crowns) # crown area in m2
ndvi_mean <- terra::global(ndvi, 'mean', na.rm = T)
evi_mean <- terra::global(evi, 'mean', na.rm = T)
gci_mean <- terra::global(gci, 'mean', na.rm = T)
lai_mean <- terra::global(lai, 'mean', na.rm = T)

canopyarea <- round(canopyarea, digits = 2)
ndvi_mean <- round(ndvi_mean, digits = 2)
evi_mean <- round(evi_mean, digits = 2)
gci_mean <- round(gci_mean, digits = 2)
lai_mean <- round(lai_mean, digits = 2)

```

## Results
```{r results,  warning=FALSE, include=TRUE, message=FALSE}
# plot indices
par(mfrow = c(2,2))
plot(ndvi, main = "ndvi")
plot(evi, main = "evi")
plot(gci, main = "gci")
plot(lai, main = "lai")

# plot tree tops
par(mfrow = c(1,2))
plot(CHM, col = height.colors(50), main = "Tree Tops")
plot(sf::st_geometry(ttops), add = TRUE, pch = 3)
plot(crowns, col = pastel.colors(200), legend = FALSE, main = "Tree Segmentation")

# save all data in one df
params_df[nrow(params_df) + 1,] <- c(EIA_name, 
                                     numtrees, 
                                     treedens, 
                                     treeheight_min, 
                                     treeheight_max, 
                                     treeheight_mean, 
                                     canopyarea$area)
params_df_indices[nrow(params_df_indices) + 1,] <- c(ndvi_mean, 
                                                     evi_mean, 
                                                     gci_mean, 
                                                     lai_mean)
kable(params_df)
kable(params_df_indices)
```
